Stack Safety is a Security Property

### 2021-04-17 tex/main.tex

\definition
For any \(\MPCT,\brk \in \MPCT^*\) with \(\head(\MPCT) = \mach,\pol,\context\),
let \(\components = \mathit{secret}(\context)\).
A system enjoys \textbf{\textit{stepwise stack confidentiality}} if,
for any \(\nach\) such that \(\mach \approx_\components \nach\) and the induced trace
\(\NPCT\) such that \(\nach,\pol,\domMap \hookrightarrow \NPCT | \brk\),
\(\MPCT \doteq \NPCT\) holds.\apt{I'm worried about quantifier order here. Surely $n$ needs to be
  universally quantified, and probably also \(\MPCT\) (?) }
\sna{Better?}

@@

In other words, stepwise stack confidentiality states that if we vary every component of the stack,
the resulting trace (starting in the same policy state) steps in lockstep with the
original trace until both have reached the break condition.
%
Note that \(\MPCT\) may contain nested calls; these must also respect the secrecy
of the components in \(\components\), and in addition their subtraces are also in
\(\MPCT^*\), and therefore confidentiality must hold on them separately. \(\NPCT\)
will contain these calls and returns as well, so its return stack pushes and pops
in lockstep until it breaks at the final return.

Finally, we formalize a version of WBCF, capturing the same informal idea
as \citet{Skorstengaard+19b} using our notion of contexts.
\definition
A system enjoys \textit{\textbf{well-bracketed control flow}} when, for every pair of reachable states
\(\mpcstatename = (\underscore,\underscore,(\underscore,\rts))\) and
\(\mpcstatename' = (\underscore,\underscore,(\underscore,\rts'))\) such that
\(\mpcstatename \stepsto \mpcstatename'\), if \(|\rts'| < |\rts|\), then
\(|\rts'| = |\rts|-1\)

@@

We first notice that, in principle, these properties appear to enforce
confidentiality outright. So we wish in particular to find an integrity
property. The intuition is that sealed memory can change, but it is
then tainted, and its original owner cannot actually read it. This
leads us to a hybrid of integrity and confidentiality.

For simplicity, our definitions return to the simple stack without sharing,
but all properties in this section can be extended to more complex settings
very naturally.

\definition
For any \(\MPCT,\brk \in \MPCT^*\), let \(\head(\MPCT) = \mach,\pol,\context\) and
suppose that \(\last(\MPCT) = \mach',\pol',\context'\) and
\((\mach',\pol',\context') \hookrightarrow \MPCT'\). Find the set of components
that changed during the call, \(\components = \prot(\context) \cap \Delta(\mach,\mach')\).
A system enjoys {\em deferred stepwise integrity} if for any \(\nach\) such that
\(\mach' \approx_{\components'} \nach\) and its induced trace
\((\nach,\pol',\context') \hookrightarrow \NPCT\), \(\MPCT' \doteq \NPCT\).

In short, if sealed components are modified during the call, the caller may not
access their values.
%
For example, consider again the integrity violation in
Figure~\ref{fig:int1}. Under a deferred discipline, the write to {\tt
  x} in instruction 11 taints the variable but does not in itself
constitute a violation, which happens only when the tainted value is
read in instruction 4. However, if this instruction added the constant
that was originally stored in {\tt x} without reading the variable,
i.e., {\tt add \$42,\%ra}, deferred integrity would not be violated.
Instead of forbidding such dangerous writes, we say,
``No reads after dangerous writes.''

@@

\definition
For any \(\MPCT,\brk \in \MPCT^*\),
let \(\head(\MPCT) = \mach,\pol,\context\).
%
Then, supposing that \(\last(\MPCT) = \mach',\pol',\context'\),
let \(\components\) be the set of protected components that changed during
the call, \(\components = \prot(\context) \cap \Delta(\mach, \mach')\),
and let \((\mach,\pol',\context') \hookrightarrow \MPCT'\).
A system enjoys {\em observational integrity} if for any \(\nach\) such that
\(\mach' \approx_{\components'} \nach\), and its induced trace
\((\nach,\pol',\context') \hookrightarrow \NPCT\), \(\obsof(\MPCT') \simeq \obsof(\NPCT)\).

@@

\begin{itemize}
\item They are observationally equivalent, \(\obsof(\MPCT) = \obsof(\NPCT)\),
  meaning that no secrets leak during the call
\item \(\brk ~ \last(\MPCT) \leftrightarrow \brk ~
\last(\NPCT)\)---if one call returns, so does the other
\item Assuming both calls return, let \(\last(\MPCT) = \mach',\pol',\context'\)
  and \(\pi_\mach(\last(\NPCT)) = \nach'\).
  Let \(\components' = \components \cap \bar{\Diamond}(\mach,\mach',\nach,\nach')\).
  And let \(\nach''\) be a state such that \(\mach' \approx_\components \nach''\).
  Then if \(\mach',\pol',\context' \hookrightarrow \MPCT'\) and
  \(\nach',\pol',\context' \hookrightarrow \NPCT'\), it must be
  that \(\obsof(\MPCT') \simeq \obsof(\NPCT')\)
\end{itemize}

>>>>

\definition
For any \(\MPCT,\brk \in \MPCT^*\) with \(\head(\MPCT) = \mach,\pol,\context\),
let \(\components = \mathit{secret}(\context)\).
A system enjoys \textbf{\textit{stepwise stack confidentiality}} if,
for any \(\nach\) such that \(\mach \approx_\components \nach\) and the induced trace
\(\NPCT\) such that \(\nach,\pol,\domMap \hookrightarrow \NPCT | \brk\),
\(\MPCT \doteq \NPCT\) holds.\apt{I'm worried about quantifier order here. Surely $n$ needs to be
  universally quantified, and probably also \(\MPCT\) (?) }
\sna{Better?\apt{only partly. Quantification of $M,\brk$ still seems wrong to me. }}

@@

In other words, stepwise stack confidentiality states that if, at the start of the call,
we vary all stack locations that are supposed to be secret from the callee
(which in our simple setting is the entirety of the stack),
the resulting trace remains in lockstep with the original trace
until either both return together or one failstops. This implies that
if the callee reads a secret location, it must failstop.\apt{right?}
%
Note that \(\MPCT\) may contain nested calls; these must also respect the secrecy
of the components in \(\components\), and in addition their subtraces are also in
\(\MPCT^*\), and therefore confidentiality must hold on them separately. \(\NPCT\)
will contain these calls and returns as well, so its return stack pushes and pops
in lockstep until it breaks at the final return.

Finally, we formalize a version of WBCF, which captures the same informal idea
as that of \citet{Skorstengaard+19b} using our context machinery.
\definition
A system enjoys \textit{\textbf{well-bracketed control flow}} when, for every pair of reachable states
\(\mpcstatename = (\underscore,\underscore,(\underscore,\rts))\) and
\(\mpcstatename' = (\underscore,\underscore,(\underscore,\rts'))\) such that
\(\mpcstatename \stepsto \mpcstatename'\)\apt{doesn't reachability of $\mpcstatename$ imply reachability of $\mpcstatename'$?}, if \(|\rts'| < |\rts|\), then
\(|\rts'| = |\rts|-1\)

@@

We first notice that, in principle\apt{??}, these properties appear to enforce
confidentiality outright. So we wish in particular to find an integrity
property. The intuition is that sealed memory can change, but it is
then tainted, and its original owner cannot actually read it. This
leads us to a hybrid of integrity and confidentiality.

%For simplicity, our definitions return to the simple stack without sharing,
%but all properties in this section can be extended to more complex settings
%very naturally.

\definition
For any \(\MPCT,\brk \in \MPCT^*\), let \(\head(\MPCT) = \mach,\pol,\context\) and
suppose that \(\last(\MPCT) = \mach',\pol',\context'\) and
\((\mach',\pol',\context') \hookrightarrow \MPCT'\). Let 
\(\components = \prot(\context) \cap \Delta(\mach,\mach')\), the
set of sealed components that changed during the call.
A system enjoys \textbf{\textit{deferred stepwise integrity}} if, for any \(\nach'\) and \(\NPCT'\)
such that \(\mach' \approx_{\components} \nach'\) and 
\((\nach',\pol',\context') \hookrightarrow \NPCT'\), \(\MPCT' \doteq \NPCT'\) holds.\apt{again, shouldn't
we quantify over all $M$?}

In short, if sealed components are modified during the call, the caller may not
access their values.\apt{This is pretty cryptic.}
%
For example, consider again the integrity violation in
Figure~\ref{fig:int1}. Under a deferred discipline, the write to {\tt
  x} in instruction 11 taints the variable but does not in itself
constitute a violation, which happens only when the tainted value is
read in instruction 4. However, if this instruction added the constant
that was originally stored in {\tt x} without reading the variable,
i.e., {\tt add \$42,\%ra}, deferred integrity would not be violated.
Instead of strictly forbidding such dangerous writes, we say,
``No reads after dangerous writes.''

@@

\definition
For any \(\MPCT,\brk \in \MPCT^*\),
let \(\head(\MPCT) = \mach,\pol,\context\).
%
Then, supposing that \(\last(\MPCT) = \mach',\pol',\context'\),
let \(\components\) be the set of protected components that changed during
the call, \(\components = \prot(\context) \cap \Delta(\mach, \mach')\),
and let \((\mach,\pol',\context') \hookrightarrow \MPCT'\).
A system enjoys {\em observational integrity} if for any \(\nach\) such that
\(\mach' \approx_{\components'} \nach\), and its induced trace
\((\nach,\pol',\context') \hookrightarrow \NPCT\), \(\obsof(\MPCT') \simeq \obsof(\NPCT)\).
\apt{Adjust quantification to match final version of deferred intergrity.}

@@

\begin{itemize}
\item They are observationally equivalent, \(\obsof(\MPCT) = \obsof(\NPCT)\),
  meaning that no secrets leak during the call
\item \(\brk ~ \last(\MPCT) \leftrightarrow \brk ~
\last(\NPCT)\)---if one call returns, so does the other
\item Assuming both calls return, let \(\last(\MPCT) = \mach',\pol',\context'\)
  and \(\pi_\mach(\last(\NPCT)) = \nach'\).
  Let \(\components' = \components \cap \bar{\Diamond}(\mach,\mach',\nach,\nach')\).
  And let \(\nach''\) be a state such that \(\mach' \approx_\components \nach''\).
  Then if \(\mach',\pol',\context' \hookrightarrow \MPCT'\) and
  \(\nach',\pol',\context' \hookrightarrow \NPCT'\), it must be
  that \(\obsof(\MPCT') \simeq \obsof(\NPCT')\)\apt{adjust quantification if needed}
\end{itemize}
